---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peakflows

<!-- badges: start -->
[![R-CMD-check](https://github.com/iluvmnms1999/peakflows/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/iluvmnms1999/peakflows/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of peakflows is to ...

## Installation

You can install the development version of peakflows from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("iluvmnms1999/peakflows")
```

## Introduction to the Data
The package was built around data provided from the USGS website (https://waterdata.usgs.gov/nwis/) which contains streamflow information for several stream gauges in the Sierra Nevadas.

<!--- Maybe we can include a table with some of the data... even a head() of what the data kind of looks like-->

```{r example}
library(peakflows)
```

## Package Functions

___ functions were created to analyse the streamflow data. 

df_peaks_all description (?)

The function df_peaks_filt... (description) 

The function peak_plot plots the peak streamflows... (description)

## Example

This is a basic example which shows you how the peakflows function works:

<!--- Maybe include a summary of our dataset/the parameters? Possibly added earlier on while explaining the dataset-->

```{r}
summary(cars) 
```

To visualize the peak flows over a specific period of time at a specific station, the peak_plot function can be used.

```{r pressure, echo = FALSE}
plot(pressure) ### create a plot using peak_flows function
```
From Figure... we can conclude...

