---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peakflows

<!-- badges: start -->
[![R-CMD-check](https://github.com/iluvmnms1999/peakflows/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/iluvmnms1999/peakflows/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction
The **peakflows** package is built around data provided from the USGS website (https://waterdata.usgs.gov/nwis/) containing streamflow information for six different stream gauges in the Sierra Nevada. Its purpose is to allow users to extract and subset peak data for specified stations and easily identify peaks representing floods, as well as differentiate between minor and major flood classifications if possible.

The peakflows package can be installed directly from the development version of peakflows from [GitHub](https://github.com/) with the following code:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("iluvmnms1999/peakflows")
```


The peakflows package can be called using the following code:

```{r setup}
library(peakflows)

```
