---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peakflows

<!-- badges: start -->
<!-- badges: end -->

## Introduction
The peakflows package is built around data provided from the USGS website (https://waterdata.usgs.gov/nwis/) which contains streamflow information for several stream gauges in the Sierra Nevadas. The purpose of the peakflows package is to allow the user to extract and subset data and easily understand when the discharged streamflow was a flood, major, or minor.

The peakflows package can be installed directly from the development version of peakflows from [GitHub](https://github.com/) with the following code:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("iluvmnms1999/peakflows")
```


The peakflows package can be called using the following code:

```{r setup}
library(peakflows)

```
