---
title: "peak-flows-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{peak-flows-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
The peakflows package is built around data provided from the USGS website (https://waterdata.usgs.gov/nwis/) which contains streamflow information for several stream gauges in the Sierra Nevadas. The purpose of the peakflows package is to allow the user to extract and subset data and easily understand when the discharged streamflow was a flood, major, or minor.

The peakflows package can be installed directly from the development version of peakflows from [GitHub](https://github.com/) with the following code:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("iluvmnms1999/peakflows")
```


The peakflows package can be called using the following code:

```{r setup}
library(peakflows)

```

## Data

The data was extracted from the USGS website (https://waterdata.usgs.gov/nwis/). The data includes information from six USGS stations from Nevada including: Carson River near Carson City (10311000), Clear Creek near Carson City (10310500), East Fork Carson River near Gardnerville (10309000), Ash Canyon Creek near Carson City (10311200), Kings Canyon Creek near Carson City (10311100), and East Fork Carson River below Markleeville (10308200).

## Subsetting the Peak Level
This function takes flooding data (peaks) from all of the stations. The user specifies the subset of flooding data that is returned by calling the df_peaks_all function and inputting either a 50, 75, or 100 as the argument. Entering a 100 returns a data frame containing peaks that are above the flooding threshold. Entering a 75 returns a data frame containing peaks that are above the flooding threshold and all peaks that are above 75% of the flooding threshold. Entering a 50 returns a data frame containing peaks that are above the flooding threshold and all peaks that are above 50% of the flooding threshold. The data frame that is returned by this function is then used for further analysis.

### Example
The following code demonstrates how the df_peaks_all function is called and the three arguments that can be used. Only portions of the df_peaks_all function is called as each dataframe contains 80, 129, and 249, respectively.

The head of df_peaks_all with an argument of 100:
```{r}
head(df_peaks_all(100))

```

The first 10 rows of df_peaks_all with argument of 75:
```{r}
df_peaks_all(75)[1:10, ]

```

## Sort by Station
Say that we are only interested at looking at the information from Kings Canyon Creek, and we want to see any floods that are at 75 percent flood level. We can filter for these specific data as well using the df_peaks_filt() function. This function is very similar to the above function but it also filters the data based on station. The station must be from one of the specifc stations in the data set and one of the specified flood levels.

### Example
Below is an example finding the first 6 peaks peaks above 75 percent flood level for Kings Canyon Creek.

```{r}
kings_can_75 <- df_peaks_filt(station = "10311100", minpeaks_spec = 75)
head(kings_can_75)

```

## Plotting the Data
The purpose of the peak_plot function is to graph the Streamflow Peaks at each station over a specific period of time. This allows the user to see whether or not there was a flood over that specific period of time at the station. The x-axis contains the range of dates provided, while the y-axis contains reflects the discharge amount. The resulting graphs contain colored dots to represent whether the peak was a flood, major, minor, or none. Red represents a flood or major amount of flood water. Yellow represents a minor amount of discharged flood water.

### Example
An example of the peak_plot function is to take a scenario where the user wants to see which stations had floods over January 2017.

```{r, warning = FALSE, fig.cap = "Figure 1: peak_flows plot for station 10311000 over January 2017", fig.height = 4, fig.width = 5}
peak_plot(station = 10311000, peaks = 100,
          beg_date = 2017010100, end_date = 2017020100)
```

Figure 1 contains the peak_plot function for the station Carson River near Carson City (10311000). This plot shows that the data contains a flood or major discharge amount of water.

```{r, warning = FALSE, fig.cap = "Figure 2: peak_flows plot for station 10311100 over January 2017", fig.height = 4, fig.width = 5}
peak_plot(station = 10311100, peaks = 100,
          beg_date = 2017010100, end_date = 2017020100)
```

Figure 2 contains the peak_plot function for the station Kings Canyon Creek near Carson City (10311100) had multiple floods or major amounts of discharge over the month of January in 2017.

## Summary

The peakflows package uses streamflow data provided from the USGS website to organize the Data through subsetting the peak level, sort by station, and plot the data over a period of time. This package allows the user to easily understand when the discharged streamflow was a flood, major, or minor. This allows for easier functionality as the user only needs to input a few values to get the output for the data.
